\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-05-31 2-the-cubble-pkg.Rmd
fix code out of text

Some readers may question why a new data structure is needed rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column. The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object. Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.

@@

```{r echo = TRUE}
# add station-based features in the nested form.
dt |> mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
```

>>>>

Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}. The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object. Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.

@@

```{r echo = TRUE}
# add station-based features in the nested form.
dt |> 
  mutate(
    fabletools::features(
      ts, tmax, list(tmax_mean = mean, tmax_var = var)
      )
  )
```

### 2022-05-31 4-1-basic.Rmd
fix code out of text

```{r echo = TRUE, warning = TRUE, message= TRUE}
cb <- as_cubble(list(spatial = lga, temporal = covid),
                key = lga, index = date, coords = c(cent_long, cent_lat))
```

@@

```{r echo = TRUE}
pair <- as_cubble(list(spatial = lga, temporal = covid),
                  key = lga, index = date, coords = c(cent_long, cent_lat),
                  output = "unmatch")

>>>>

```{r echo = TRUE, warning = TRUE, message= TRUE}
cb <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat)
  )
```

@@

```{r echo = TRUE}
pair <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat),
  output = "unmatch"
  )

### 2022-05-31 jss/cubble-jss.tex
fix code out of text

The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}. The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.

@@

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 1,830 x 10
  id            lat  long  elev name         wmo_id date        prcp  tmax  tmin
  <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <date>     <dbl> <dbl> <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-01     0  31.9  15.3
2 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-02     0  24.9  16.4
3 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-03     6  23.2  13  
4 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-04     0  28.4  12.4
5 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- climate_flat |>
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested |>
+   mutate(rain_day = sum(ts$prcp != 0))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts       rain_day
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <int>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>      104
2 ASN00010311 -31.9  117. 179   york            94623 <tibble>       89
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>       90
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>      106
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>       63
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested |> face_temporal()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long |> face_spatial()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\begin{CodeInput}
R> identical(cubble_nested, cubble_back)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold <- cubble_long |> unfold(long, lat)
R> cubble_unfold
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold |> face_spatial() |> face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # example with a tsibble created from climate_flat
R> raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
R> dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # add station-based features in the nested form.
R> dt |> mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts       tmax_mean tmax_var
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>       <dbl>    <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> covid |> head(5)
\end{CodeInput}
\begin{CodeOutput}
# A tsibble: 5 x 5 [1D]
# Key:       lga [1]
# Groups:    lga, source [1]
  date       lga        source                            n roll_mean
  <date>     <chr>      <chr>                         <int>     <dbl>
1 2022-01-01 Alpine (S) Contact with a confirmed case     1        NA
2 2022-01-02 Alpine (S) Contact with a confirmed case     2        NA
3 2022-01-03 Alpine (S) Contact with a confirmed case     4        NA
4 2022-01-04 Alpine (S) Contact with a confirmed case     4        NA
5 2022-01-05 Alpine (S) Contact with a confirmed case     2        NA
\end{CodeOutput}
\begin{CodeInput}
R> lga |> head(5)
\end{CodeInput}
\begin{CodeOutput}
Simple feature collection with 5 features and 7 fields
Geometry type: MULTIPOLYGON
Dimension:     XY
Bounding box:  xmin: 142.3535 ymin: -38.67876 xmax: 147.3909 ymax: -36.39269
Geodetic CRS:  WGS 84
    lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
132         20110     Alpine (S)               2        Victoria     4788.1568
133         20260    Ararat (RC)               2        Victoria     4211.1171
134         20570   Ballarat (C)               2        Victoria      739.0321
135         20660    Banyule (C)               2        Victoria       62.5402
136         20740 Bass Coast (S)               2        Victoria      865.8095
    cent_long  cent_lat                       geometry
132  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
133  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
134  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
135  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
136  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(list(spatial = lga, temporal = covid),
+                 key = lga, index = date, coords = c(cent_long, cent_lat))
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have corresponding spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have corresponding temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pair <- as_cubble(list(spatial = lga, temporal = covid),
+                   key = lga, index = date, coords = c(cent_long, cent_lat),
+                   output = "unmatch")
R> 
R> pair
\end{CodeInput}
\begin{CodeOutput}
$paired
# A tibble: 2 x 2
  spatial             temporal    
  <chr>               <chr>       
1 Kingston (C) (Vic.) Kingston (C)
2 Latrobe (C) (Vic.)  Latrobe (C)

@@

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [8]: nested form
# bbox:     [144.52, -37.73, 146.06, -36.55]
# temporal: date [date], matched_var [dbl]
  id          name                  lat  long type   dist group ts       n_match
  <chr>       <chr>               <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
8 ASN00088051 redesdale           -37.0  145. clim~  1.84     1 <tibble>      18
\end{CodeOutput}
\end{CodeChunk}

>>>>

The main functions in the cubble package are \code{as_cubble(), face_spatial(), face_temporal(), and, unfold()}.

The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and, \code{unfold()}. The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.

@@

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 1,830 x 10
  id         lat  long  elev name  wmo_id date        prcp  tmax  tmin
  <chr>    <dbl> <dbl> <dbl> <chr>  <dbl> <date>     <dbl> <dbl> <dbl>
1 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-01     0  31.9  15.3
2 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-02     0  24.9  16.4
3 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-03     6  23.2  13  
4 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-04     0  28.4  12.4
5 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- climate_flat |>
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested |>
+   mutate(rain_day = sum(ts$prcp != 0))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts       rain_day
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>      <int>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble>      104
2 ASN00010311 -31.9  117. 179   york           94623 <tibble>       89
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble>       90
4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tibble>      106
5 ASN00015131 -17.6  134. 220   elliott        94236 <tibble>       63
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested |> face_temporal()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long |> face_spatial()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
\end{CodeOutput}
\begin{CodeInput}
R> identical(cubble_nested, cubble_back)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold <- cubble_long |> unfold(long, lat)
R> cubble_unfold
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold |> face_spatial() |> face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # example with a tsibble created from climate_flat
R> raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
R> dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # add station-based features in the nested form.
R> dt |> 
+   mutate(
+     fabletools::features(
+       ts, tmax, list(tmax_mean = mean, tmax_var = var)
+       )
+   )
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id          lat  long  elev name  wmo_id ts       tmax_mean tmax_var
  <chr>     <dbl> <dbl> <dbl> <chr>  <dbl> <list>       <dbl>    <dbl>
1 ASN00009~ -31.9  116.  15.4 pert~  94610 <tbl_ts>      25.7    38.6 
2 ASN00010~ -31.9  117. 179   york   94623 <tbl_ts>      26.2    51.1 
3 ASN00010~ -32.9  117. 338   narr~  94627 <tbl_ts>      23.7    45.4 
4 ASN00014~ -12.4  131.  30.4 darw~  94120 <tbl_ts>      33.1     3.02
5 ASN00015~ -17.6  134. 220   elli~  94236 <tbl_ts>      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> covid |> head(5)
\end{CodeInput}
\begin{CodeOutput}
# A tsibble: 5 x 5 [1D]
# Key:       lga [1]
# Groups:    lga, source [1]
  date       lga        source                            n roll_mean
  <date>     <chr>      <chr>                         <int>     <dbl>
1 2022-01-01 Alpine (S) Contact with a confirmed case     1        NA
2 2022-01-02 Alpine (S) Contact with a confirmed case     2        NA
3 2022-01-03 Alpine (S) Contact with a confirmed case     4        NA
4 2022-01-04 Alpine (S) Contact with a confirmed case     4        NA
5 2022-01-05 Alpine (S) Contact with a confirmed case     2        NA
\end{CodeOutput}
\begin{CodeInput}
R> lga |> head(5)
\end{CodeInput}
\begin{CodeOutput}
Simple feature collection with 5 features and 7 fields
Geometry type: MULTIPOLYGON
Dimension:     XY
Bounding box:  xmin: 142.3535 ymin: -38.67876 xmax: 147.3909 ymax: -36.39269
Geodetic CRS:  WGS 84
    lga_code_2018            lga state_code_2016 state_name_2016
132         20110     Alpine (S)               2        Victoria
133         20260    Ararat (RC)               2        Victoria
134         20570   Ballarat (C)               2        Victoria
135         20660    Banyule (C)               2        Victoria
136         20740 Bass Coast (S)               2        Victoria
    areasqkm_2018 cent_long  cent_lat                       geometry
132     4788.1568  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
133     4211.1171  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
134      739.0321  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
135       62.5402  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
136      865.8095  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat)
+   )
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pair <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat),
+   output = "unmatch"
+   )
R> 
R> pair
\end{CodeInput}
\begin{CodeOutput}
$paired
# A tibble: 2 x 2
  spatial             temporal    
  <chr>               <chr>       
1 Kingston (C) (Vic.) Kingston (C)
2 Latrobe (C) (Vic.)  Latrobe (C)

@@

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [8]: nested form
# bbox:     [144.52, -37.73, 146.06, -36.55]
# temporal: date [date], matched_var [dbl]
  id          name        lat  long type   dist group ts       n_match
  <chr>       <chr>     <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
1 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
3 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
4 ASN00082170 benalla ~ -36.6  146. clim~  8.54    10 <tibble>      21
5 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
6 ASN00086038 essendon~ -37.7  145. clim~  6.17     6 <tibble>      19
7 406213      CAMPASPE~ -37.0  145. river  1.84     1 <tibble>      18
8 ASN00088051 redesdale -37.0  145. clim~  1.84     1 <tibble>      18
\end{CodeOutput}
\end{CodeChunk}
